Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2014-01-17 paper/ms.tex
paper updated. New plot added to give support to the exclusion of the third family of models whre deltaM>1.0

\title[Halo mass and occupation fraction of LAEs at$z=3.1$]{Mass and
  occupation fraction of dark matter halos hosting Lyman-$\alpha$
  emitters at $z\sim 3$}      
\author[~J.~E. Forero-Romero and ~J.~E. Mejia-Restrepo]{
\parbox[t]{\textwidth}{\raggedright 
  Jaime E. Forero-Romero$^{1}$ and
  Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

Recently \cite{Yamada2012} observed a wide area of $2.4$ deg$^{2}$
under homogeneous instrumental and data reduction conditions. This data
set is constructed from 12 different sub-fields that allows us to use
clustering statistics and cosmic variance to constraint the mass and
occupation fraction of halos hosting LAEs.

@@

Figure \ref{fig:correlation_parameters} shows the results for the
best estimates of $\theta_{0}$-$\beta$  used in the ACF
parameterization.  Blue circles with error bars represent the results
from the mocks and the horizontal line the observational results of
\cite{Ouchi2010} over a field with average number density where the
$\beta$ parameter was fixed in the fit. From this Figure we observe
that there are models with too large values of $\theta_0$ and $\beta$
that can be ruled out as inconsistent with the angular correlation
function.

@@

Comparing Figure
\ref{fig:restriction_mock_and_f_occ_corr} with Figure
\ref{fig:restriction_mock} we see that the new constraint disfavors
almost all models with $M_{\rm max}>10^{12}$. The only exception is 
a narrow range of models with $M_{\rm min}=10^{10.9}$. However, we
have verified that these models are only marginally consistent with
the ACF constraints in the $\theta_0$-$\beta$ plane.

From this comparison we conclude that clustering information can
constraint the values of $M_{\rm max}$, but not the occupation
fraction.

\begin{figure}
\begin{center}
\includegraphics[width=1.0\linewidth,angle=0]{power_law_correlation_degree2.pdf}  
\end{center}
\caption{Values for the free parameters $\theta_{0}$ and $\beta$
in the fitting formula (Eq. \ref{eq:fitting}) for the angular
correlation function. Blue dots correspond to simulations and the
green line to observations by \citet{Ouchi2008,Ouchi2010}. The error
bars in the theoretical data correspond to the quadratic average of
the fitting uncertainties for each mock
survey \label{fig:correlation_parameters}}  
\end{figure}

@@

{\bf The models in the third family are barely consistent with the
  constraints from the ACF. They have values of $\beta$ close to $1.0$
  (instead of the observational value of $0.8$) and $\theta_0$ on the
  range of $13^{\prime}$ (the mean observational value is close to
  $8^{\prime}$) but are considered
  consistent because their large error bars barely overlap with the
  observational expectation. These 
  models they have another particular feature: their minimum mass is
  exactly $M_{\rm min}=10^{10.9}\hMsun$. Because these models are
  are barely compatible with observations and are a minority of the
  consistent models, we exclude them from the rest of the discussion.}

@@

{\bf A fourth caveat is the comparing results for the number density
  statistics and the ACF from surveys with different EW cuts. In this
  sense we highlight againt that the two filters and selection
  criteria used by \citep{Ouchi2008} and \citep{Yamada2012} are only
  slightly different. The median number density in the general fields by
  \citep{Yamada2012} is $0.20$ arcmin$^{-2}$, while in
  \citep{Ouchi2008} is $0.099\pm0.005$ arcmin$^{-2}$. This gives us
  confidence that the two different samples can give a self-consistent picture
  within a factor of $\sim2$ for our studies.}

@@

{\bf We find that the information for the number density constrains
  the minium mass to the range
  $10^{10}\hMsun < M_{\rm min} < 10^{11.5}\hMsun$ while the clustering
  constrains the upper mass $M_{\rm    max} < 10^{12}\hMsun$. However,
  this information insufficient to impose a tight constraint  on the escape
  fraction. Additionaly, we find that the mass range defined as
$\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm min}$ is $\Delta
M\leq 1.0$, with a large fraction of models with $\Delta M\leq 0.5$.}

{\bf Our approach has allowed the discovery of possible models that have
been overloooked so far in the literature. A central role in these new
models is played by the wide range of escape fraction values,
something unexpected from previous studies \citep[i.e.][]{Gawiser2007,Ouchi2010}  that unanimously favor low values on the range of $10^{-2}$. The largest
innovation in our analysis is the construction of mock obsevations
that directly explore the impact of cosmic variance.}

Nevertheless, all the halo mass ranges deduced previous analysis are a
subset of the models we find in this paper.  All of our
preferred models support the notion that the most massive halos at
$z=3.1$ do not host detectable LAEs. {\bf Furthermore, we claim that the
mass range for halos hosting LAEs is narrow, $\Delta M\leq 1.0$, with
a majority of models with $\Delta M\leq 0.5$.}

@@

In summary, we show that {\bf the currently available spatial
  information of LAEs cannot put a tight constraint on their
  occupation fraction of dark matter halos}. This information can
confidently be used to establish that these galaxies are hosted in low
mass halos with masses spanning less than one order of magnitude. {\bf
  From the point of view of Dark Energy surveys such as the Hobby-Eberly
  Telescope Dark Enery Experiment (HETDEX) the
  narrow mass range for LAEs can be seen as an advantage because its
  simplicity in their modeling aiming at recoverin cosmological
  parameters.} We foresee that the observations with new instruments
(such as the Multi Unit Spectroscopic Explorer (MUSE), Hyper
SuprimeCam and HETDEX) covering larger fields and a wider range of
luminosities will be key in reducing cosmic variance and imposing
tighter constraints on the properties of dark matter halos hosting
LAEs. At the same time, additional modeling for \ly radiation transfer
is needed to put tighter constraints on  occupation fraction in high
redshift halos, also paying attention to other physical phenomena,
such as the stochasticity \citep{ForeroRomero2013} in the star
formation process, which might play a role in inducing detection
biases in high redshift LAEs.

>>>>

\title[Halo mass and occupation fraction of LAEs at$z=3.1$]{Mass and
  occupation fraction of dark matter halos hosting Lyman-$\alpha$
  emitters at $z\sim 3$}      
\author[~J.~E. Forero-Romero and ~J.~E. Mejia-Restrepo]{
\parbox[t]{\textwidth}{\raggedright 
  Jaime E. Forero-Romero$^{1}$ and
  Julian E. Mej\'ia-Restrepo$^{2,3}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile\\
$^3$FACom-Instituto de Física-FCEN, Universidad de Antioquia, Calle 70 No. 52-21, Medellín, Colombia}

@@

Recently \cite{Yamada2012} observed a wide area of $2.4$ deg$^{2}$
under homogeneous instrumental and data reduction conditions. This data
set is constructed from 12 different sub-fields that allows us to use
clustering statistics and cosmic variance to constrain the mass and
occupation fraction of halos hosting LAEs.

@@

The left panel of the figure \ref{fig:correlation_parameters} shows the results for the
best estimates of $\theta_{0}$-$\beta$  used in the ACF
parameterization.  Blue circles with error bars represent the results
from the mocks and the horizontal line the observational results of
\cite{Ouchi2010} over a field with average number density where the
$\beta$ parameter was fixed in the fit. From this Figure we observe
that there are models with too large values of $\theta_0$ and $\beta$
that can be ruled out as inconsistent with the angular correlation
function.

@@

Comparing Figure
\ref{fig:restriction_mock_and_f_occ_corr} with Figure
\ref{fig:restriction_mock} we see that the new constraint disfavors
almost all models with $M_{\rm max}>10^{12}$. The only exception is 
a narrow range of models with $M_{\rm min}=10^{10.9}$. However, in the right panel of figure  \ref{fig:correlation_parameters}   we can observe that these models are only 
marginally consistent with the ACF constraints in  $\theta_0$.

From this comparison we conclude that clustering information can
constrain the values of $M_{\rm max}$, but not the occupation
fraction.

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{Fig7_corr_params.pdf}  
\includegraphics[width=0.46\linewidth,angle=0]{Fig7_mmin_vs_theta.pdf} 
\end{center}
\caption{Left panel. Values for the free parameters $\theta_{0}$ and $\beta$
in the fitting formula (Eq. \ref{eq:fitting}) for the angular
correlation function. Blue dots correspond to simulations and the
green line to observations by \citet{Ouchi2008,Ouchi2010}. The error
bars in the theoretical data correspond to the quadratic average of
the fitting uncertainties for each mock survey. Right panel. Values of $\theta_{0}$  and $\log_{10}(M_{min}/M_{\odot} h^{-1})$. Blue and green point represents the models where  $\log_{10}(M_{max}/M_{\odot} h^{-1})<12.0$ and  $\log_{10}(M_{min}/M_{\odot} h^{-1})>12.0$ respectively. .  The horizontal gray band represents the observational contraints on $theta_{0}$ stablished by   \citet{Ouchi2008,Ouchi2010}. Green points has been displaced by $0.01dex$ of its original value to avoid overlapping between blue and green points.} \label{fig:correlation_parameters}  
\end{figure*}

@@

{\bf The models in the third family are barely consistent with the
  constraints from the ACF. These 
  models they have another particular feature: their minimum mass is
  exactly $M_{\rm min}=10^{10.9}\hMsun$ m and $M_{\rm max}>10^{12.0}\hMsun$ . They have values of $\beta$ close to $1.0$
  (instead of the observational value of $0.8$) and $\theta_0$ on the
  range of $13^{\prime}$ (the mean observational value is close to
  $8^{\prime}$) but are considered
  consistent because their large error bars barely overlap with the
  observational expectation (see figure \ref{fig:correlation_parameters})  .  Because these models are
  are barely compatible with observations and are a minority of the
  consistent models, we exclude them from the rest of the discussion.}

@@

{\bf A fourth caveat is the comparing results for the number density
  statistics and the ACF from surveys with different EW cuts. In this
  sense we highlight again that the two filters and selection
  criteria used by \citep{Ouchi2008} and \citep{Yamada2012} are only
  slightly different. The median number density in the general fields by
  \citep{Yamada2012} is $0.20$ arcmin$^{-2}$, while in
  \citep{Ouchi2008} is $0.099\pm0.005$ arcmin$^{-2}$. This gives us
  confidence that the two different samples can give a self-consistent picture
  within a factor of $\sim2$ for our studies.}

@@

{\bf We find that the information for the number density constrains
  the minium mass to the range
  $10^{10}\hMsun < M_{\rm min} < 10^{11.5}\hMsun$ while the clustering
  constrains the upper mass $M_{\rm    max} < 10^{12}\hMsun$. However,
  this information is insufficient to impose a tight constraint  on the escape
  fraction. Additionaly, we find that the mass range defined as
$\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm min}$ is $\Delta
M\leq 1.0$, with a large fraction of models with $\Delta M\leq 0.5$.}

{\bf Our approach has allowed the discovery of possible models that have
been overlooked so far in the literature. A central role in these new
models is played by the wide range of escape fraction values,
something unexpected from previous studies \citep[i.e.][]{Gawiser2007,Ouchi2010}  that unanimously favor low values on the range of $10^{-2}$. The largest
innovation in our analysis is the construction of mock obsevations
that directly explore the impact of cosmic variance.}

Nevertheless, all the halo mass ranges deduced in previous analysis are a
subset of the models we find in this paper.  All of our
preferred models support the notion that the most massive halos at
$z=3.1$ do not host detectable LAEs. {\bf Furthermore, we claim that the
mass range for halos hosting LAEs is narrow, $\Delta M\leq 1.0$, with
a majority of models with $\Delta M\leq 0.5$.}

@@

In summary, we show that {\bf the currently available spatial
  information of LAEs cannot put a tight constraint on their
  occupation fraction of dark matter halos}. This information can
confidently be used to establish that these galaxies are hosted in low
mass halos with masses spanning less than one order of magnitude. {\bf
  From the point of view of Dark Energy surveys such as the Hobby-Eberly
  Telescope Dark Enery Experiment (HETDEX) the
  narrow mass range for LAEs can be seen as an advantage because its
  simplicity in their modeling aiming at recovering cosmological
  parameters.} We foresee that the observations with new instruments
(such as the Multi Unit Spectroscopic Explorer (MUSE), Hyper
SuprimeCam and HETDEX) covering larger fields and a wider range of
luminosities will be key in reducing cosmic variance and imposing
tighter constraints on the properties of dark matter halos hosting
LAEs. At the same time, additional modeling for \ly radiation transfer
is needed to put tighter constraints on  occupation fraction in high
redshift halos, also paying attention to other physical phenomena,
such as the stochasticity \citep{ForeroRomero2013} in the star
formation process, which might play a role in inducing detection
biases in high redshift LAEs.
